Numerical simulation and entropy dissipative cure of the carbuncle instability for the shallow water circular hydraulic jump

### 2020-11-16 numerics_paper/version2/numerical_chj.tex
update paper

\subsection{Our contribution}
\begin{itemize}
\item We present a new problem where the cc instability occurs. In our opinion, this is a more challenging 
  problem than the standard bow problem since we expect the solution to be non-steady. 
  We aim to preserve instabilities that are inherently present in the CHJ for certain regimes. And, 
  at the same time, we aim to prevent the cc instability. 
\item We explore the use of entropy dissipation and stability as cure of the cc instability for the SWEs. 
\item We propose specific methods to blend the highly dissipative and robust Rusanov's RP and the 
  accurate 3-wave Roe's solver for the SWEs. In particular we propose methods that induce dissipation of entropy
  near the shocks and other methods that guarantee entropy stability. 
\item We demonstrate that these schemes prevent the formation of cc instabilities not only 
  in the well-known bow shock problem, but also in the CHJ while preserving other important features of 
  the solution. 
\end{itemize}

@@

\subsection{Semi-analytical steady solution under rotational symmetry}

By assuming rotational symmetry in \eqref{eq:sw}, one obtains the
system (in one spatial variable)
\begin{subequations} \label{eq:rsw}
\begin{align}
    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
\end{align}
\end{subequations}
Steady-state solutions of \eqref{eq:rsw} satisfy
\begin{subequations}
\begin{align}
    rhu & = C \\
    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
\end{align}
\end{subequations}
for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
We see that two types of steady profiles exist, depending on whether the flow
is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.

@@

\subsection{Formation of the CHJ}

@@

The accuracy of a given first- or second-order method depends heavily on the Riemann solver. 
In particular, the Rusanov's Riemann solver is highly dissipative. In contrast, the well-known 
Roe's Riemann solver produces more accurate solutions. 
%
As we discussed in the introduction, finite volume methods with 
Roe's Riemann solvers are prone to the carbuncle instability for the CHJ. 
In contrast, if we use the Rusanov's 
Riemann solver, no carbuncles appear; however, many other important and physical features of 
the solution are dissipated, unless the grid is highly refined. 
Our main objective is to study the CHJ and to propose a Riemann solver
that is carbuncle free but do not dissipate important features of the solution. 
To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
Our blending criteria is based on the residual of the entropy (in)equality. 
Similar ideas have been explored before to solve the Euler equations XX 
and the shallow water equations XX. In these references, the authors use different 
blending criteria.

\subsection{Lax-Wendroff-LeVeque method}
Let us now review the second-order correction to method \eqref{first-order_FV} implemented in 
Clawpack \cite{clawpack} and PyClaw \cite{pyclaw-sisc}.
We refer to \cite{leveque1997wave, leveque2002finite} for a detailed description of the method. 
We consider only a one-dimensional problem; in the 
numerical experiments in \S\ref{sec:num} we use second-order Strang's splitting \cite{strang1968construction} 
to extend the method to multiple dimensions.

@@

\subsection{Entropy stable Riemann solver}
The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
measurement of the entropy residual. The objective is to induce dissipation of entropy 
near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct} 
or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable. 
In this section, we consider the one-dimensional problem and increase the numerical dissipation 
to guarantee entropy stability for \eqref{first-order_via_fluct}. 
To this end we follow the work by Tadmor, see for instance \cite{tadmor1987numerical, tadmor2003entropy} 
and references therein.

@@

\clearpage
\section{Numerical results}\label{sec:num}
Description of the methods. We consider either the first-order method XX or the second-order Lax-Wendroff
type method XX. For each of these methods we use one of the following three Riemann solvers: 
\begin{itemize}
  \item LLF: Rusanov's or Local Lax Friederichs 2-wave Riemann solver reviewed in \S XX.
  \item Roe: standard Roe's average based 3-wave Riemann solver reviewed in \S XX. 
  \item Roe-fix: Roe's solver with entropy fix based on XX. 
  \item EV:  entropy dissipative 3-wave Riemann solver described in \S XX.
\end{itemize}

\subsection{Dam break problem on a dry bed}
We start with the one-dimensional dam break problem on a dry bed. 
This problem is a canonical example that demonstrates the `entropy glitch' of Roe's solver 
at transonic rarefactions. Since the baseline Roe's solver we use (see \S XX) does not contain 
an entropy fix, it is important to demonstrate that the blended Riemann solver fixes the entropy glitch, 
at least for this benchmark. 
We follow the setup in XX. The domain is given by XX, and the initial condition is XX and XX. 
The exact solution, which can be found in XX and references therein, is

XX

We use g=1 and solve the problem up to T=10. In Figure XX we show the solution using method XX 
with Roe's solver, Rusanov's solver and the blended solver. The entropy glitch, which is manifested 
as a non-physical show at x=XX, is clearly shown with Roe's solver. Using Rusanov's solver and the 
blended solver fix the entropy glitch. Using the blended solver leads to more accurate results. 
In Table XX, we summarize the results of a convergence test.

\subsection{Dam break problem on a wet bed}
We consider a one-dimensional dam break problem on a wet domain. 
We follow the setup in XX. 
The domain is $\Omega=(0,L)$ with $L=10$, the initial condition is given by 
$hu(x,0)=0$ and 
\begin{align*}
  h(x,0) = 
  \begin{cases}
    h_l & \mbox{ if } x\in(0,x_0], \\
    h_r & \mbox{ if } x\in(x_0,L),
  \end{cases}
\end{align*}
with $x_0=5$, $h_l=0.005$ and $h_r=0.001$. 
The exact solution, which can be found in XX and references therein, is given by 
\begin{align*}
  h(x,t) = 
  \begin{cases}
    h_l, \\
    \frac{4}{9g}\left(\sqrt{gh_l}-\frac{x-x_0}{2t}\right)^2, \\
    \frac{c_m^2}{g}, \\
    h_r, 
  \end{cases}
\quad 
  u(x,t) = 
  \begin{cases}
    0, &\mbox{ if } x\leq x_A(t), \\
    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
    2(\sqrt{gh_l}-c_m), & \mbox{ if } x_B(t)<x\leq x_C(t), \\
    0, &\mbox{ if } x\leq x_C(t),
  \end{cases}
\end{align*}
where $x_A(t)=x_0-t\sqrt{gh_l}$, $x_B(t)=x_0+t\left(2\sqrt{gh_l}-3c_m\right)$, 
$x_C(t)=x_0+t\frac{2c_m^2\left(\sqrt{gh_l}-c_m\right)}{c_m^2-gh_r}$ and 
$c_m$ is the solution of 
$-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
We use $g=1$ and solve the problem up to the final time $T=5$ considering the 2nd-order Lax-Wendroff 
type method XX with the LLF, Roe and EV Riemann solvers. The solution with different refinement levels 
and each Riemann solver is shown in Figure XX. In Table XX, we summarize the results of a convergence test. 
Since the solution is non-smooth, we expect no more than first order convergence rates. Note that the 
results with the entropy dissipative solver EV are comparable against the results using Roe's solver. 
That is, inducing entropy dissipation via the EV Riemann solver does not degrade the high-order accuracy 
properties of the Roe's solver. 
In contrast, the accuracy and convergence rates using the LLF solver are clearly degraded.

\subsection{The circular hydraulic jump in one dimension}

\subsection{Steady outflow}
Let us consider a problem similar to that in the previous section, but with an outflow boundary condition. 
That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition is 
$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition is $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$, 
and the right boundary condition is set to outflow, see the details in XX. 
If the flow is supercritical ($|F|>1$), the exact solution converges to a 
steady state profile given by the solution of the ODE XX. 
Depending on the initial condition, shocks might develop before the steady state is reached. 
Nevertheless, the steady state profile is smooth. 
In Figure XX, we show the solution at different times based on the secnd-order method XX with 
the EV Riemann solver. Additionally, in Table XX we summarize the result of a convergence study 
based on the first- and second-order methods XX and XX, respectively. 
We consider the Riemann solver in \S XX with Ri=0, Ri=1 and Ri computed via XX.

\subsection{The circular hydraulic jump in two dimensions}
We consider two regimes. In the first regime, the physical instabilities are mild. 
By increasing the strength of the jump and other parameters, we consider a second 
and more strongly unstable regime. In both cases, we generate the CHJ as described in 
Section XX; that is, we impose supersonic inflow boundary conditions for the jet and 
subsonic outflow boundary conditions.

We consider two types of domain and two corresponding computational grids. 
The two types of domains are two-dimensional squares and circular domains. 
For both cases we use structured grids. See an example of these domains and grids in Figures XX and XX, 
respectively.

\subsection{Regime 1: midly unstable flow}

Show the results with Roe's and Rusanov's solver. 
Show the results with the three proposed methodologies.

\subsection{Regime 2: strongly unstable flow}

Show the results with Roe's and Rusanov's solver. 
Show the results with the three proposed methodologies.

>>>>

\subsection{Our contribution}
\begin{itemize}
\item We present a new problem where the cc instability occurs. In our opinion, this is a more challenging 
  problem than the standard bow shock problem since we expect the solution to be non-steady. 
  We aim to preserve instabilities that are inherently present in the CHJ for certain regimes. And, 
  at the same time, we aim to prevent the cc instability. 
\item We explore the use of entropy dissipation and entropy stability as cure of the cc instability for the SWEs. 
\item We propose a specific method to blend the highly dissipative and robust Rusanov's RP and the 
  accurate 3-wave Roe's solver for the SWEs. 
\item We demonstrate that this scheme prevents the formation of cc instabilities not only 
  in the well-known bow shock problem, but also in the CHJ while preserving other important features of 
  the solution. 
\end{itemize}

@@

\subsection{Semi-analytical steady solution under rotational symmetry}\label{sec:steady_chj}

By assuming rotational symmetry in \eqref{eq:sw}, one obtains the
system (in one spatial variable)
\begin{subequations} \label{eq:rsw}
\begin{align}
    (rh)_t + (rhu)_r & = 0 \label{mass1} \\
    (rhu)_t + (rhu^2)_r + r \left(\frac{1}{2}gh^2\right)_r = 0. \label{mom1}
\end{align}
\end{subequations}
Steady-state solutions of \eqref{eq:rsw} satisfy
\begin{subequations}\label{steady}
\begin{align}
    rhu & = C \\
    h'(r) & = \frac{h}{\frac{g}{\beta^2} r^3 h^3 -r} = \frac{h}{r} \cdot \frac{F^2}{1-F^2} \label{eq:dh}
\end{align}
\end{subequations}
for some $C$ independent of $r$.  Here $F=|u|/\sqrt{gh}$ is the Froude number.
We see that two types of steady profiles exist, depending on whether the flow
is subcritical ($|F|<1$) or supercritical ($|F|>1$).  No smooth steady solution can
include both regimes, since the right hand side of \eqref{eq:dh} blows up when $F=1$.

@@

\subsection{Formation of the CHJ}\label{sec:formation}

@@

The accuracy of a given first- or second-order method depends heavily on the Riemann solver. 
In particular, the Rusanov's Riemann solver is highly dissipative. In contrast, the well-known 
Roe's Riemann solver produces more accurate solutions. 
%
As we discussed in the introduction, finite volume methods with 
Roe's Riemann solvers are prone to the carbuncle instability for the CHJ. 
In contrast, if we use the Rusanov's 
Riemann solver, no carbuncles appear; however, many other important and physical features of 
the solution are dissipated, unless the grid is highly refined. 
Our main objective is to study the CHJ and to propose a Riemann solver
that is carbuncle free but do not dissipate important features of the solution. 
To do that we propose a Riemann solver that blends Rusanov's and Roe's solvers. 
Our blending criteria is based on the residual of the entropy (in)equality,
which induces dissipation of entropy near discontinuities. 
Blending these Riemann solvers has been proposed before
to solve the Euler equations, see e.g.
\cite{nishikawa2008very,wang2016developing,jaisankar2007diffusion,ohwada2018simple,deng2019new,ray2013entropy},
the shallow water equations, see e.g. \cite{bader2014carbuncle,kemm2014note},
and even the Navier-Stokes equations, see e.g. \cite{nishikawa2008very,ohwada2018simple}.
In these references, the authors use blending criteria different than ours.
Nevertheless, fixing the carbuncle instability by imposing entropy stability is not new.
For example, in \cite{ismail2009affordable,ismail2009proposed},
the authors developed carbuncle free solvers for the Euler equations
via dissipation of entropy.

\subsection{Lax-Wendroff-LeVeque method}\label{sec:lax-wendroff-leveque}
Let us now review the second-order correction to method \eqref{first-order_FV} implemented in 
Clawpack \cite{clawpack} and PyClaw \cite{pyclaw-sisc}.
We refer to \cite{leveque1997wave, leveque2002finite} for a detailed description of the method. 
We consider only a one-dimensional problem; in the 
numerical experiments in \S\ref{sec:num} we use second-order Strang's splitting \cite{strang1968construction} 
to extend the method to multiple dimensions.

@@

\subsection{Entropy stable Riemann solver}\label{sec:entropy_stable}
The Riemann solver presented in the previous section blends Rusanov's and Roe's solver based on a
measurement of the entropy residual. The objective is to induce dissipation of entropy 
near discontinuities. However, we have no guarantee that methods \eqref{first-order_via_fluct} 
or \eqref{second-order_via_fluct} based on this Riemann solver are (locally) entropy stable. 
In this section, we consider the one-dimensional problem and increase the numerical dissipation 
to guarantee entropy stability for \eqref{first-order_via_fluct}. 
To this end we follow the work by Tadmor, see for instance \cite{tadmor1987numerical, tadmor2003entropy} 
and references therein.

@@

\begin{remark}[About the effect of $\lambda_{i\pm 1/2}^{\min}$ for Roe's solver.]
  If we set $R_{i\pm 1/2}=0, ~\forall i$ in \eqref{lambda_p}, then the Riemann solver becomes the Roe's solver.
  In this case, it is well known that transonic rarefactions develop non-physical shocks,
  a phenomenon known as `entropy-glitch'. To fix this problem, many entropy fixes to the Roe's
  solver have been developed. Enhancing the dissipation via \eqref{lambda_min} behaves as an entropy fix for
  the Roe's solver, which we demonstrate via a numerical experiment in \S \ref{sec:rp_dry_bed}. 
\end{remark}

\clearpage
\section{Numerical results}\label{sec:num}
In this section we present one- and two-dimensional numerical experiments to demonstrate the behavior of the
blended Riemann solver from \S\ref{sec:blended_rs} with the extra entropy dissipation from \S\ref{sec:entropy_stable}. 
We compare the behavior of the blended solver against the standard Roe's and Rusanov's solvers.
In most of the experiments we use these Riemann solvers with the Lax-Wendroff-LeVeque method reviewed in \S\ref{sec:lax-wendroff-leveque}.
When the exact solution is available, we report a convergence test based on the $L^1$-error for the water height
\begin{align*}
  E_1=\sum_i|K_i|~|h_i-h^e(\bfx_i)|,
\end{align*}
where $|K_i|$ is the length or area of cell $i$,
$h_i$ is the cell average of the water height at cell $i$,
and $h^e(\bfx_i)$ is the exact water height evaluated at the center of cell $i$.

We consider a total of six problems. We start with a one-dimensional Riemann problem over a dry bed. 
For this problem we use the first-order method \eqref{first-order_via_fluct},
which helps preventing violations of positivity. Although, only the Rusanov's solver guarantees
positivity, we do not get negative values for the water height with this method and the other Riemann solvers.
In \S \ref{sec:steady_outflow} we consider a one-dimensional problem with a smooth steady state solution. We observe the expected
second-order accuracy of method \eqref{second-order_via_fluct} with Roe's and the blended solvers. Using
the Rusanov's Riemann solver degrades the accuracy to first-order. The overall accuracy of the blended solver
is not degraded since the extra dissipation is not applied in smooth regions. 
In \S\ref{sec:rp_wet_bed} and \S\ref{sec:1D_chj} we consider one-dimensional problems with strong shocks. We observe similar accuracy with the
blended solver than with Roe's solver; that is, inducing entropy dissipation near the shocks do not degrade the accuracy
of the underlying Roe's solver. This extra dissipation, however, prevents the formation of carbuncles in other experiments,
which we demonstrate in \S \ref{sec:bow_shock} and \S\ref{sec:2D_chj}. 
The main focus of this work is in the formation of carbuncle instabilities in the two-dimensional CHJ. We present
an extensive set of experiments about this problem in \S \ref{sec:2D_chj}.
We consider not only Roe's, Rusanov's and the blended solvers, but also the HLLE solver and a solver proposed in \cite{kemm2014note},
which is designed to remove the carbuncle instabilities in the shallow water equations.

\subsection{Dam break problem on a dry bed}\label{sec:rp_dry_bed}
We start with the one-dimensional dam break problem on a dry bed. 
This problem is a canonical example that demonstrates the `entropy glitch' of Roe's solver 
at transonic rarefactions. Since the baseline Roe's solver we use (see \S\ref{sec:roe}) does not contain 
an entropy fix, it is important to demonstrate that the blended Riemann solver fixes the entropy glitch.
We follow the setup in \cite[\S 4.1.2]{delestre2013swashes}.
The domain is given by $\Omega=(0,L=10)$, and the initial condition is $hu(x,0)=0$ and
\begin{align*}
  h(x,0)=
  \begin{cases}
    h_l, & \mbox{ if }, x\leq x_0=5 \\
    h_r, &\mbox{ otherwise},
  \end{cases}
\end{align*}
with $h_l=5\times 10^{-3}$ and $h_r=0$. In our experiments we use $h_r=1\times 10^{-15}$ to avoid division by zero.
The exact solution, which can also be found in \cite{delestre2013swashes} and references therein, is 
\begin{align*}
  h(x,t) = 
  \begin{cases}
    h_l, \\
    \frac{4}{9g}\left(\sqrt{gh_l}-\frac{x-x_0}{2t}\right)^2, \\
    0,
  \end{cases}
\quad 
  u(x,t) = 
  \begin{cases}
    0, &\mbox{ if } x\leq x_A(t), \\
    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
    0, &\mbox{ if } x_B(t)\leq x,
  \end{cases}
\end{align*}
with $x_A(t)=x_0-t\sqrt{gh_l}$ and $x_B(t)=x_0+2t\sqrt{gh_l}$.
We use $g=1$ and solve the problem up to $t=10$.
In figures \ref{fig:rp_dry_bed_roe}-\ref{fig:rp_dry_bed_blended},
we show the solution using method \eqref{first-order_via_fluct} 
with Roe's solver, Rusanov's solver and the blended solver, respectively.
The entropy glitch, which is manifested as a non-physical shock at $x=x_0$,
is present with Roe's solver. Using Rusanov's and the blended solvers
fixes the entropy glitch.
In Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test. 
Note that using the blended solver leads to more accurate results.

\begin{figure}[!h]
  {\scriptsize
    \subfloat[Roe's Riemann solver.\label{fig:rp_dry_bed_roe}]{
      \includegraphics[scale=0.36]{figures/rp_dry_roe.png}}
    ~
    \subfloat[Rusanov's Riemann solver.\label{fig:rp_dry_bed_llf}]{
        \includegraphics[scale=0.36]{figures/rp_dry_llf.png}}

    \subfloat[Blended Riemann solver. \label{fig:rp_dry_bed_blended}]{
      \includegraphics[scale=0.36]{figures/rp_dry_blended.png}}
    ~
    \subfloat[Roe's Riemann solver with $\lambda_{i\pm 1/2}^{\min}$.\label{fig:rp_dry_bed_roe_with_lmin}]{
      \includegraphics[scale=0.36]{figures/rp_dry_roe_with_lmin.png}}    
  }
  \caption{
    Dam break problem over a dry bed using method \eqref{first-order_via_fluct} with different
    Riemann solvers. We show the numerical solution and the exact solution (in dashed black) at $t=10$.
    We consider different refinements with
    (cyan) $\Delta x=1/400$, (red) $\Delta x=1/800$, (blue) $\Delta x=1/1600$.
    \label{fig:rp_dry_bed}}
\end{figure}

For this particular problem, using method \eqref{first-order_via_fluct} with the blended Riemann solver
leads to $\lambda_{i\pm 1/2}^{\min}=0$ for all the experiments, but this is not true in general.
Finally, let us demonstrate the effect of $\lambda_{i\pm 1/2}^{\min}$ by imposing $R_{i}=0$ in \eqref{lambda_p},
which is equivalent to using Roe's solver with extra dissipation given by $\lambda_{i\pm 1/2}^{\min}$.
In Figure \ref{fig:rp_dry_bed_roe_with_lmin}, we show the solution using different refinements,
and in the last column of Table \ref{table:rp_dry_bed}, we summarize the results of a convergence test.
Note that the entropy glitch is highly reduced. 
To remove completely the entropy glitch we could add high-order entropy dissipation following \cite{tadmor2003entropy}
and references therein.

\begin{table}[!ht]\scriptsize
  \begin{center}
    \begin{tabular}{||c||c|c||c|c||c|c||c|c||} \hline
      & \multicolumn{2}{c||}{Roe's solver}  
      &\multicolumn{2}{c||}{Rusanov's solver} 
      &\multicolumn{2}{c||}{Blended solver}
      &\multicolumn{2}{c||}{Roe's with $\lambda_{i\pm 1/2}^{\min}$} \\ \cline{2-9}
      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
      1/50   & 6.21E-04 &  --  & 6.95E-04 &   -- & 5.10E-04 &  --  & 5.95E-04 & --   \\
      1/100  & 4.70E-04 & 0.40 & 5.27E-04 & 0.40 & 3.53E-04 & 0.52 & 4.06E-04 & 0.55 \\
      1/200  & 3.91E-04 & 0.26 & 3.88E-04 & 0.44 & 2.38E-04 & 0.56 & 2.83E-04 & 0.52 \\
      1/400  & 3.16E-04 & 0.30 & 2.64E-04 & 0.55 & 1.51E-04 & 0.65 & 1.89E-04 & 0.58 \\
      1/800  & 2.35E-04 & 0.42 & 1.67E-04 & 0.66 & 9.43E-05 & 0.68 & 1.20E-04 & 0.65 \\
      1/1600 & 2.00E-04 & 0.23 & 1.01E-04 & 0.72 & 5.70E-05 & 0.72 & 7.42E-05 & 0.68 \\ \hline
    \end{tabular}
    \caption{Grid convergence study for the dam break problem over a dry bed
      using method \eqref{first-order_via_fluct} with different Riemann solvers.\label{table:rp_dry_bed}}
  \end{center}
\end{table}

\subsection{Dam break problem on a wet bed}\label{sec:rp_wet_bed}
We consider now a one-dimensional dam break problem on a wet domain. 
We follow the setup in \cite[\S 4.1.1]{delestre2013swashes}. 
The domain is $\Omega=(0,L)$ with $L=10$, the initial condition is given by 
$hu(x,0)=0$ and 
\begin{align*}
  h(x,0) = 
  \begin{cases}
    h_l & \mbox{ if } x\in(0,x_0], \\
    h_r & \mbox{ if } x\in(x_0,L),
  \end{cases}
\end{align*}
with $x_0=5$, $h_l=0.005$ and $h_r=0.001$. 
The exact solution, which can be found in \cite{delestre2013swashes} and references therein,
is given by 
\begin{align*}
  h(x,t) = 
  \begin{cases}
    h_l, \\
    \frac{4}{9g}\left(\sqrt{gh_l}-\frac{x-x_0}{2t}\right)^2, \\
    \frac{c_m^2}{g}, \\
    h_r, 
  \end{cases}
\quad 
  u(x,t) = 
  \begin{cases}
    0, &\mbox{ if } x\leq x_A(t), \\
    \frac{2}{3}\left(\sqrt{gh_l}+\frac{x-x_0}{t}\right), & \mbox{ if } x_A(t) < x\leq x_B(t), \\
    2(\sqrt{gh_l}-c_m), & \mbox{ if } x_B(t)<x\leq x_C(t), \\
    0, &\mbox{ if } x\leq x_C(t),
  \end{cases}
\end{align*}
where $x_A(t)=x_0-t\sqrt{gh_l}$, $x_B(t)=x_0+t\left(2\sqrt{gh_l}-3c_m\right)$, 
$x_C(t)=x_0+t\frac{2c_m^2\left(\sqrt{gh_l}-c_m\right)}{c_m^2-gh_r}$ and 
$c_m$ is the solution of 
$-8gh_rc_m^2\left(\sqrt{gh_l}-c_m\right)^2+\left(c_m^2-gh_r\right)^2\left(c_m^2+gh_r\right)=0$.
We use $g=1$ and solve the problem up to the final time $t=5$ considering method \eqref{second-order_via_fluct}
with Roe's, Rusanov's and the blended solvers. The solution with different refinement levels 
and each Riemann solver is shown in Figure \ref{fig:rp_wet_bed}.
In Table \ref{table:rp_wet_bed}, we summarize the results of a convergence test. 
Since the solution is non-smooth, we expect no more than first order convergence rates. Note that the 
results with the entropy dissipative blended solver are comparable against the results using Roe's solver. 
That is, inducing entropy dissipation via the blended Riemann solver does not degrade the high-order
accuracy properties of the Roe's solver. 
In contrast, the accuracy and convergence rates using Rusanov's solver are clearly degraded.

\begin{figure}[!h]
  {\scriptsize
    \subfloat[Roe's Riemann solver.]{
      \includegraphics[scale=0.29]{figures/rp_wet_roe.png}}
    \quad
    \subfloat[Rusanov's Riemann solver.]{
        \includegraphics[scale=0.29]{figures/rp_wet_llf.png}}
    \quad
    \subfloat[Blended Riemann solver.]{
      \includegraphics[scale=0.29]{figures/rp_wet_blended.png}}
  }
  \caption{
    Dam break problem over a wet bed using method \eqref{second-order_via_fluct} with different
    Riemann solvers. We show the numerical solution and the exact solution (in dashed black) at $t=10$.
    We consider different refinements with
    (cyan) $\Delta x=1/400$, (red) $\Delta x=1/800$, (blue) $\Delta x=1/1600$.
    \label{fig:rp_wet_bed}}
\end{figure}

\begin{table}[!ht]\scriptsize
  \begin{center}
    \begin{tabular}{||c||c|c||c|c||c|c||} \hline
      & \multicolumn{2}{c||}{Roe's solver}  
      &\multicolumn{2}{c||}{Rusanov's solver} 
      &\multicolumn{2}{c||}{Blended solver} \\ \cline{2-7}
      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
      1/50   & 4.22E-04 &  --  & 5.73E-04 &  --  & 4.34E-04 & --   \\
      1/100  & 2.00E-04 & 1.07 & 3.35E-04 & 0.78 & 2.08E-04 & 1.05 \\
      1/200  & 1.11E-04 & 0.85 & 2.01E-04 & 0.74 & 1.13E-04 & 0.88 \\
      1/400  & 5.05E-05 & 1.13 & 1.18E-04 & 0.77 & 5.26E-05 & 1.10 \\
      1/800  & 2.60E-05 & 0.95 & 7.08E-05 & 0.74 & 2.73E-05 & 0.94 \\
      1/1600 & 1.29E-05 & 1.01 & 3.79E-05 & 0.90 & 1.32E-05 & 1.04 \\ \hline
    \end{tabular}
    \caption{Grid convergence study for the dam break problem over a wet bed
      using method \eqref{second-order_via_fluct} with different Riemann solvers.\label{table:rp_wet_bed}}
  \end{center}
\end{table}

\subsection{The circular hydraulic jump in one-dimension}\label{sec:1D_chj}
In this section we consider for the first time a CHJ. We start with a one-dimensional problem
embedded into the two-dimensional domain $\Omega=(0.1,1)\times(0,0.1)$.
A semi-analytical solution at steady state is available as described in \S \ref{sec:steady_chj};
therefore, we can perform a convergence test. 
We consider the initial condition $h(x,t=0)=0.1$, $hu(x,t=0)=0$ and $hv(x,t=0)=0$.
At the left boundary, we impose $h(0.1,t)=0.3$, $hu(0.1,t)=0.225$ and $hv(0.1,t)=0$.
At the right boundary we impose $h(1,t)=0.37387387318873766$, $hu(1,t)=\beta/h(1,t)$ and $hv(1,t)=0$ where $\beta=0.0225$.
The values on the right boundary are chosen so that the CHJ is located at $x=0.3$, as described in \S \ref{sec:formation}.
%
In the first row of Figure \ref{fig:rp_wet_bed}, we show (slices of) the solution (at $y=0.05$)
at different times using the second-order method \eqref{second-order_via_fluct} with Roe's, Rusanov's and the blended Riemann solvers. 
In the second row of the same figure, we plot $R_i$ \eqref{Ri}, which is a (normalized) measurement of the
entropy residual. Note that $R_i$ accurately locates the shocks and is (nearly) zero in smooth regions. 
In Table \ref{table:chj_1D}, we summarize the results of a convergence study. 
We remark that the convergence with Roe's solver is erratic.
Moreover, for the finer grids, the solution appears to become worse.
Using Rusanov's and the blended Riemann solvers delivers cleaner convergence rates. 
In \cite[\S 5]{kemm2018heuristical} and \cite{bader2014carbuncle}, the authors consider one-dimensional problems embedded into
two-dimensional domains and observe different instabilities for the
Euler and the shallow water equations, respectively.
The authors attributed the problem to a manifestation of the carbuncle instability
that lead to an unstable location of the steady shock with certain Riemann solvers, including Roe's solver.

\begin{figure}[!h]
  {\scriptsize
    \includegraphics[scale=0.29]{figures/chj_1D_t0p5.png}
    \qquad
    \includegraphics[scale=0.29]{figures/chj_1D_t1p3.png}
    \qquad
    \includegraphics[scale=0.29]{figures/chj_1D_t10p0.png}

    \vspace{15pt}
    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t0p5.png}
    \qquad
    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t1p3.png}
    \qquad
    \includegraphics[scale=0.29]{figures/chj_1D_Ri_t10p0.png}
  }
  \caption{
    One-dimensional CHJ using method \eqref{second-order_via_fluct} with 
    (cyan) Roe's, (red) Rusanov's and (blue) the blended Riemann solvers.
    In the second row we plot $R_i$, which is given by \eqref{Ri}.
    In all simulations we consider $\Delta x=1/400$.
    \label{fig:1D_chj}}
\end{figure}

\begin{table}[!ht]\scriptsize
  \begin{center}
    \begin{tabular}{||c||c|c||c|c||c|c||} \hline
      & \multicolumn{2}{c||}{Roe's solver}  
      &\multicolumn{2}{c||}{Rusanov's solver} 
      &\multicolumn{2}{c||}{Blended solver} \\ \cline{2-7}
      $\Delta x$ & $E_1$ & rate & $E_1$ & rate & $E_1$ & rate \\ \hline
      1/50  & 1.56E-04 &  --   & 8.80E-04 &  --  & 3.56E-04 &  --  \\
      1/100 & 1.06E-04 & 0.55  & 5.09E-04 & 0.79 & 1.95E-04 & 0.86 \\
      1/200 & 7.77E-05 & 0.45  & 2.96E-04 & 0.78 & 1.07E-04 & 0.87 \\
      1/400 & 7.50E-06 & 3.37  & 1.49E-04 & 0.99 & 3.68E-05 & 1.53 \\
      1/800 & 9.07E-06 & -0.27 & 7.98E-05 & 0.90 & 1.67E-05 & 1.14 \\ \hline
    \end{tabular}
    \caption{Grid convergence study for the one-dimensional CHJ at $t=10$
      using method \eqref{second-order_via_fluct} with different Riemann solvers.\label{table:chj_1D}}
  \end{center}
\end{table}

\subsection{Steady outflow}\label{sec:steady_outflow}
Let us consider a problem similar to that in the previous section, but with an outflow boundary condition. 
That is, we consider the one-dimensional domain $\Omega=(0.1,1)$, the initial condition 
$h(x,t=0)=0.1$, $hu(x,t=0)=0$, the left boundary condition $h(0.1,t)=0.3$, $hu(0.1,t)=0.75$, 
and set the right boundary condition to outflow.
We refer to \cite[\S 21.8.5]{leveque2002finite} for details about the outflow boundary condition. 
If the flow is supercritical ($|F|>1$ where $F$ is the Froude number), the exact solution converges to a 
steady state profile given by the solution of \eqref{steady}. 
Depending on the initial condition, shocks might develop before the steady state is reached. 
Nevertheless, the steady state profile is smooth. 
In Figure \ref{fig:steady_outflow}, we show the solution and $R_i$ at different times using
the second-order method \eqref{second-order_via_fluct}
with Roe's, Rusanov's and the blended Riemann solvers.
Additionally, in Table \ref{table:steady_outflow}, we summarize the results of a convergence study 
based on methods \eqref{first-order_via_fluct} and
\eqref{second-order_via_fluct}, using the same Riemann solvers.

\begin{figure}[!h]
  {\scriptsize
    \includegraphics[scale=0.29]{figures/outflow_1D_t0p5.png}
    \qquad
    \includegraphics[scale=0.29]{figures/outflow_1D_t1p3.png}
    \qquad
    \includegraphics[scale=0.29]{figures/outflow_1D_t10p0.png}

    \vspace{15pt}
    \includegraphics[scale=0.29]{figures/outflow_1D_Ri_t0p5.png}
    \qquad
    \includegraphics[scale=0.29]{figures/outflow_1D_Ri_t1p3.png}
    \qquad
    \includegraphics[scale=0.29]{figures/outflow_1D_Ri_t10p0.png}
  }
  \caption{
    Steady outflow problem using method \eqref{second-order_via_fluct} with 
    (cyan) Roe's, (red) Rusanov's and (dashed blue) the blended Riemann solvers.
    In the second row we plot $R_i$, which is given by \eqref{Ri}.
    In all simulations we consider $\Delta x=1/400$.
    \label{fig:steady_outflow}}
\end{figure}
